---
title: "DA2"
format: 
  html:
    toc: true
    embed-resources: true
    code-fold: true
    code-line-numbers: true
    code-summary: "SHOW ME THE CODE!"
    code-tools: true
editor: visual
---

```{r setup}
#| message: false
#| warning: false
library(lme4)
library(tidyverse)
library(AICcmodavg)
library(gtsummary)
library(dplyr)
library(e1071)
```

```{r}
# read data
demo1 <- read.csv("./STARD Demo 1.csv")
demo2 <- read.csv("./STARD Demo 2.csv")
starQids <- read.csv("./STARD QIDS.csv")
```

# Part I

## a)

```{r}
# merge demograph datasets
demo <- merge(demo1, demo2, by = "ID")
```

```{r}
# clean data
colMeans(is.na(demo))
d2 <- demo[complete.cases(demo),]
colMeans(is.na(d2))
colMeans(is.na(starQids))
sq <- starQids[complete.cases(starQids),]
colMeans(is.na(sq))

```

## b)

```{r}
# remove anything with .x
qids.3 <- starQids[-grep("\\.", starQids$WEEK), ]

# count of subjects in qids.3
length(unique(qids.3$ID))

# Get maximum week per subject
max_weeks <- aggregate(WEEK ~ ID, data = qids.3, max)

# Convert max week to number of visits
max_weeks$Visits <- (max_weeks$WEEK / 2) + 1

# Filter subjects with at least 3 visits
qids.3 <- qids.3[qids.3$ID %in% max_weeks$ID[max_weeks$Visits >= 3], ]

# Create Table 1: Count of subjects by number of visits
table_1 <- table(max_weeks$Visits[max_weeks$Visits >= 3])
table_1

# count of subjects in qids.3 after filtering subjects with >=3 visits
length(unique(qids.3$ID))
```

## c)

```{r}
# Create a vector of unique IDs from QIDS.3
valid_ids <- unique(qids.3$ID)

# Filter the demo dataset to include only these subjects
demo_filtered <- demo %>% filter(ID %in% valid_ids)

# Check sample size
length(unique(demo_filtered$ID))

# Sample size is one less than qids.3

# Check for IDs in QIDS.3 that are missing in demo_filtered
#missing_in_demo <- setdiff(unique(qids.3$ID), demo_filtered$ID)
#missing_in_demo

#Turns out ID 2172 is missing in the demo dataset so we remove it from our dataset


# Replace invalid values (-5 for categorical, <0 for numeric)
demo_filtered <- demo_filtered %>% filter(ID != 2172)

# Replace both -3 and -5 with NA for variables where these values represent missing/invalid data
demo_filtered$DEP[demo_filtered$DEP == -3 | demo_filtered$DEP == -5] <- NA
demo_filtered$BIP[demo_filtered$BIP == -3 | demo_filtered$BIP == -5] <- NA
demo_filtered$ALCOHOL[demo_filtered$ALCOHOL == -3 | demo_filtered$ALCOHOL == -5] <- NA
demo_filtered$DRUG_PHX[demo_filtered$DRUG_PHX == -3 | demo_filtered$DRUG_PHX == -5] <- NA
demo_filtered$SUIC_PHX[demo_filtered$SUIC_PHX == -3 | demo_filtered$SUIC_PHX == -5] <- NA

# Replace all negative values with NA for epino, episode_date and dage
demo_filtered$HSPNC[demo_filtered$HSPNC < 0] <- NA
demo_filtered$epino[demo_filtered$epino < 0] <- NA
demo_filtered$dage[demo_filtered$dage < 0] <- NA


# Since all episode_date values are negative, we will make them positive
demo_filtered$episode_date <- abs(demo_filtered$episode_date)
```

```{r}
# Histogram of 'dage'
ggplot(demo_filtered, aes(x = dage)) + 
  geom_histogram(bins = 30, fill = "blue", alpha = 0.7) +
  labs(title = "Histogram of Age (dage)", x = "Age", y = "Frequency")

# Histogram of 'epino'
ggplot(demo_filtered, aes(x = epino)) + 
  geom_histogram(bins = 30, fill = "green", alpha = 0.7) +
  labs(title = "Histogram of Episodes (epino)", x = "Number of Episodes", y = "Frequency")

# Histogram of 'episode_date'
ggplot(demo_filtered, aes(x = episode_date)) + 
  geom_histogram(bins = 30, fill = "blue", alpha = 0.7) + 
  labs(title = "Distribution of Episode Date", x = "Episode Date", y = "Frequency")


# Calculate skewness while ignoring NAs
skewness(demo_filtered$dage, na.rm = TRUE)
skewness(demo_filtered$epino, na.rm = TRUE)
skewness(demo_filtered$episode_date, na.rm = TRUE)

```

```{r}
# Classify race categories # THIS NEEDS TO BE SPECIFIED IN THE REPORT
demo_filtered <- demo_filtered %>%
  mutate(
    # White race classification
    Race_White = ifelse(WHITE == 1, "White", NA),
    
    # Black race classification
    Race_Black = ifelse(BLACK == 1 & !(WHITE == 1), "Black", NA),
    
    # Asian race classification, only for those who did not select Black
    Race_Asian = ifelse(ASIAN == 1 & !(WHITE == 1 | BLACK == 1), "Asian", NA),
    
    # Other race classification: for subjects who did not select White, Black, or Asian
    Race_Other = ifelse(
      (HSPNC == 1 | AMIND == 1 | HAWAI == 1) & !(WHITE == 1 | BLACK == 1 | ASIAN == 1),
      "Other",
      NA
    )
  )

# Summary for continuous variables (Median and IQR for skewed data)
summary_continuous <- demo_filtered %>%
  summarise(
    dage_median = median(dage, na.rm = TRUE),
    dage_IQR = IQR(dage, na.rm = TRUE),
    episode_date_median = median(episode_date, na.rm = TRUE),
    episode_date_IQR = IQR(episode_date, na.rm = TRUE),
    epino_median = median(epino, na.rm = TRUE),
    epino_IQR = IQR(epino, na.rm = TRUE)
  )

# Summary for categorical variables (N and percentages for each category)
summary_categorical <- demo_filtered %>%
  summarise(
    SEX_Female_n = sum(SEX == 2, na.rm = TRUE),  # Count for Female
    SEX_Female_pct = sum(SEX == 2, na.rm = TRUE) / n() * 100,  # Percentage for Female
    SEX_Male_n = sum(SEX == 1, na.rm = TRUE),  # Count for Male
    SEX_Male_pct = sum(SEX == 1, na.rm = TRUE) / n() * 100,  # Percentage for Male

    DEP_No_n = sum(DEP == 0, na.rm = TRUE),  # Count for No depression history
    DEP_No_pct = sum(DEP == 0, na.rm = TRUE) / n() * 100,  # Percentage for No depression history
    DEP_Yes_n = sum(DEP == 1, na.rm = TRUE),  # Count for Yes depression history
    DEP_Yes_pct = sum(DEP == 1, na.rm = TRUE) / n() * 100,  # Percentage for Yes depression history

    ALCOHOL_No_n = sum(ALCOHOL == 0, na.rm = TRUE),  # Count for No alcohol use
    ALCOHOL_No_pct = sum(ALCOHOL == 0, na.rm = TRUE) / n() * 100,  # Percentage for No alcohol use
    ALCOHOL_Yes_n = sum(ALCOHOL == 1, na.rm = TRUE),  # Count for Yes alcohol use
    ALCOHOL_Yes_pct = sum(ALCOHOL == 1, na.rm = TRUE) / n() * 100,  # Percentage for Yes alcohol use

    DRUG_PHX_No_n = sum(DRUG_PHX == 0, na.rm = TRUE),  # Count for No drug use
    DRUG_PHX_No_pct = sum(DRUG_PHX == 0, na.rm = TRUE) / n() * 100,  # Percentage for No drug use
    DRUG_PHX_Yes_n = sum(DRUG_PHX == 1, na.rm = TRUE),  # Count for Yes drug use
    DRUG_PHX_Yes_pct = sum(DRUG_PHX == 1, na.rm = TRUE) / n() * 100,  # Percentage for Yes drug use

    SUIC_PHX_No_n = sum(SUIC_PHX == 0, na.rm = TRUE),  # Count for No suicidal history
    SUIC_PHX_No_pct = sum(SUIC_PHX == 0, na.rm = TRUE) / n() * 100,  # Percentage for No suicidal history
    SUIC_PHX_Yes_n = sum(SUIC_PHX == 1, na.rm = TRUE),  # Count for Yes suicidal history
    SUIC_PHX_Yes_pct = sum(SUIC_PHX == 1, na.rm = TRUE) / n() * 100,  # Percentage for Yes suicidal history

    # Race Summary
    Race_White_n = sum(Race_White == "White", na.rm = TRUE),  # Count for White
    Race_White_pct = sum(Race_White == "White", na.rm = TRUE) / n() * 100,  # Percentage for White

    Race_Black_n = sum(Race_Black == "Black", na.rm = TRUE),  # Count for Black
    Race_Black_pct = sum(Race_Black == "Black", na.rm = TRUE) / n() * 100,  # Percentage for Black

    Race_Asian_n = sum(Race_Asian == "Asian", na.rm = TRUE),  # Count for Asian
    Race_Asian_pct = sum(Race_Asian == "Asian", na.rm = TRUE) / n() * 100,  # Percentage for Asian

    Race_Other_n = sum(Race_Other == "Other", na.rm = TRUE),  # Count for Other
    Race_Other_pct = sum(Race_Other == "Other", na.rm = TRUE) / n() * 100  # Percentage for Other
  )

# Combine continuous and categorical summaries into a final table
summary_table <- bind_cols(summary_continuous, summary_categorical)


print(summary_table)


### This table can be used to make a nicer table in word/pdf for the report. Look at the STARD paper for inspo. page 4 of 13.

```

## d)

In this analysis, subjects were allowed to select multiple races, and it was necessary to classify them into broader categories for reporting purposes. The following classification rules were applied to categorize race:

1.  **White**: Subjects who identified as **White** were classified as **White**.
2.  **Black**: Subjects who identified as **Black** and did **not** identify as **White** were classified as **Black**. This ensures that individuals who selected both **White** and **Black** are not mistakenly categorized as **Black**.
3.  **Asian**: Subjects who identified as **Asian** and did **not** identify as either **Black** or **White** were classified as **Asian**.
4.  **Other**: Subjects who selected races such as **Hispanic**, **AMIND**, or **HAWAI**, and did **not** select **White**, **Black**, or **Asian**, were classified as **Other**.

This classification approach ensures that individuals who selected multiple races are categorized based on their most specific race (e.g., **Black** if **Black** and **White** were selected). Additionally, the **Other** category includes subjects who selected races outside the main classifications, ensuring that their race is still represented in the dataset without misclassification.

## e)

```{r}
# Create the new composite variable 'Any.Fam.Psyc.Hist'
demo_filtered <- demo_filtered %>%
  mutate(
    Any_Fam_Psyc_Hist = ifelse(DEP == 1 | BIP == 1 | ALCOHOL == 1 | DRUG_PHX == 1 | SUIC_PHX == 1, 1, 0)
  )

```

# Part II

## a)

```{r}
library(ggplot2)
library(dplyr)

set.seed(111)

# Randomly select 100 unique subjects
rand_ids <- sample(unique(starQids$ID), size = 100, replace = FALSE)

# Filter dataset to only include the selected subjects
rand_samp <- starQids %>% filter(ID %in% rand_ids)

# Create the spaghetti plot
g1 <- ggplot(data = rand_samp, aes(x = WEEK, y = QSTOT, group = ID, color = factor(ID))) +
  geom_line(alpha = 0.5) +  # Reduce opacity for clarity
  scale_y_continuous(name = "QIDS Score") +
  scale_x_continuous(name = "Week") +
  labs(title = "Spaghetti Plot of QIDS Scores Over Time", color = "Subject ID") +
  theme_bw() +  # Use a clean theme
  theme(legend.position = "none")  # Hide legend for readability

# Print the plot
print(g1)

```

## b)

```{r}

# Jitter week variable
starQids$jweek <- jitter(starQids$WEEK)

# Create the loess curve plot
g_loess <- ggplot(data = starQids, aes(x = jweek, y = QSTOT)) +
  geom_jitter(alpha = 0.3, color = "blue") +  # Jittered points for visibility
  scale_y_continuous(name = "QIDS Score") +
  scale_x_continuous(name = "Week", breaks = 0:16) +
  stat_smooth(method = "loess", se = FALSE, color = "red", size = 1.2) +  # Loess smoothing
  labs(title = "Loess Curve of QIDS Scores Across All Subjects") +
  theme_minimal()

# Print the plot
print(g_loess)


```

## c)

```{r}
# Merge gender info with QIDS dataset
sub_cols <- demo_filtered[, c("SEX", "ID")]
sq_gen <- merge(x = starQids, y = sub_cols, by = "ID")

# Check for missing values
colSums(is.na(sq_gen))

# Jitter week values for better visualization
sq_gen$jweek <- jitter(sq_gen$WEEK)

# Create faceted plot by gender
g_gender <- ggplot(data = sq_gen, aes(x = jweek, y = QSTOT, color = factor(SEX))) +
  geom_jitter(alpha = 0.3) +  # Add jitter for better visibility
  scale_y_continuous(name = "QIDS Score") +
  scale_x_continuous(name = "Week", breaks = 0:16) +
  stat_smooth(method = "loess", se = FALSE, size = 1.2) +  # Loess curve for each facet
  facet_wrap(~SEX, labeller = labeller(SEX = c("1" = "Male", "2" = "Female"))) +  
  labs(title = "QIDS Scores Over Time by Gender", color = "Gender") +
  theme_minimal()

# Print the plot
print(g_gender)

```

## d)

```{r}
# Create a single Race_Collapsed variable
demo_filtered <- demo_filtered %>%
  mutate(
    Race_Collapsed = case_when(
      Race_White == "White" ~ "White",
      Race_Black == "Black" ~ "Black",
      Race_Asian == "Asian" ~ "Asian",
      Race_Other == "Other" ~ "Other",
      TRUE ~ NA_character_  # Assign NA for any missing values
    )
  )

# Merge race info with QIDS dataset
sub_eth <- demo_filtered[, c("ID", "Race_Collapsed")]
sq_eth <- merge(starQids, sub_eth, by = "ID")

# Jitter week values for better visualization
sq_eth$jweek <- jitter(sq_eth$WEEK)

# Create faceted plot by ethnicity
g_ethn <- ggplot(data = sq_eth, aes(x = jweek, y = QSTOT, color = factor(Race_Collapsed))) +
  geom_jitter(alpha = 0.3) +  # Reduce overplotting
  scale_y_continuous(name = "QIDS Score") +
  scale_x_continuous(name = "Week", breaks = 0:16) +
  stat_smooth(method = "loess", se = FALSE, size = 1.2) +  # Add loess curve for each category
  facet_wrap(~Race_Collapsed) +  
  labs(title = "QIDS Scores Over Time by Ethnicity", color = "Ethnicity") +
  theme_minimal()

print(g_ethn)


```

## e)

```{r}
# Merge 'epino' data with QIDS dataset
sub_episodes <- demo_filtered[, c("ID", "epino")]
sq_epi <- merge(starQids, sub_episodes, by = "ID")

# Ensure 'epino' has valid values before binning
sq_epi <- sq_epi %>% filter(!is.na(epino))  # Remove NAs if needed

# Bin 'epino' into 3 groups using cut_number
sq_epi$epi2 <- cut_number(sq_epi$epino, n = 3, labels = c("Low", "Medium", "High"))

# Create faceted plot by episode binning
g_episode <- ggplot(data = sq_epi, aes(x = WEEK, y = QSTOT)) +
  geom_jitter(alpha = 0.3, color = "blue") +  # Add jitter for better visualization
  scale_y_continuous(name = "QIDS Score") +
  scale_x_continuous(name = "Week", breaks = 0:16) +
  stat_smooth(method = "loess", se = FALSE, color = "red", size = 1.2) +  # Loess curve
  facet_wrap(~epi2) +  # Facet by episode binning
  labs(title = "QIDS Scores Over Time by Episode Count", x = "Week", y = "QIDS Score") +
  theme_minimal()

print(g_episode)

```

## f)

```{r}

# Merge gender (`SEX`) with episode-binned data (`sq_epi`)
sq_epi_sex <- merge(sq_epi, sub_cols, by = "ID")

# Check for missing values
colSums(is.na(sq_epi_sex))

# Create faceted plot by gender with separate loess curves for each episode bin
g_gender_epi <- ggplot(data = sq_epi_sex, aes(x = WEEK, y = QSTOT, color = epi2)) +
  geom_jitter(alpha = 0.3) +  # Reduce overplotting
  scale_y_continuous(name = "QIDS Score") +
  scale_x_continuous(name = "Week", breaks = 0:16) +
  stat_smooth(method = "loess", se = FALSE, size = 1) +  # Loess curve for each bin
  facet_wrap(. ~ SEX, labeller = labeller(SEX = c("1" = "Male", "2" = "Female"))) +  # Gender labels
  labs(
    title = "QIDS Scores Over Time by Gender & Episode Count",
    color = "Episode Bins"
  ) +
  theme_minimal()

# Print the plot
print(g_gender_epi)


```

# Part III

## a)

```{r}
full_data <- merge(sq, demo_filtered, by="ID")
lmer.1 <- lmer(QSTOT ~ 1 + WEEK + (1 + WEEK | ID), 
               data = full_data, 
               REML = F,
               control = lmerControl(optimizer = "Nelder_Mead"))
summary(lmer.1)
```

You can use this for the report:

**Interpreting Your Model Output (`lmer.1`)**

**Key results:** - **Fixed Effects:** - **Intercept (13.54)** → The estimated **QIDS score at Week 0** is **13.54**. - **WEEK (-0.59)** → On average, **QIDS scores decrease by 0.59 per week**. - **t-value for WEEK (-70.57)** is large, meaning **WEEK has a strong effect**.

-   **Random Effects:**
    -   **Intercept variance (12.31, SD = 3.51)** → Some subjects start with much higher or lower QIDS scores.
    -   **Slope variance for WEEK (0.078, SD = 0.28)** → Subjects vary slightly in their rate of improvement.
    -   **Residual variance (10.85, SD = 3.29)** → This represents within-subject variation in QIDS scores.
    -   **Intercept-Slope Correlation (0.02)** → Essentially **no correlation** between baseline QIDS score and rate of improvement.
-   **Model Fit:**
    -   **AIC = 93562.5, BIC = 93608.8** → These will be used to compare with the expanded model.

# b)

```{r}
full_data$Race_Collapsed <- as.factor(full_data$Race_Collapsed)

lmer.2 <- lmer(QSTOT ~ 1 + WEEK + SEX + Race_Collapsed + dage + epino + Any_Fam_Psyc_Hist + (1 + WEEK | ID), 
               data = full_data, 
               REML = F,
               control = lmerControl(optimizer = "Nelder_Mead"))
summary(lmer.2)
```

You can use this for the report:

**Model Summary**: - **Fixed Effects:** - **Intercept (10.65)**: This is the estimated **QIDS score** at **Week 0** (baseline).\
- **WEEK (-0.60)**: On average, the **QIDS score decreases by 0.60 per week**, suggesting that depression severity decreases over time. - **SEX (0.53)**: Males have **0.53 higher QIDS scores** than females, on average. - **Race_Collapsed**:\
- **Black (3.37)**: Black subjects tend to have **3.37 higher QIDS scores** than the reference category (likely "White").\
- **Other (3.89)**: Subjects from other racial backgrounds tend to have **3.89 higher QIDS scores** than the reference category.\
- **White (1.81)**: White subjects tend to have **1.81 higher QIDS scores** compared to the reference group. - **dage (-0.01)**: For every one-year increase in **age**, the **QIDS score decreases slightly by 0.01**. - **epino (0.01)**: **Epilepsy history** has a **small positive effect** on QIDS scores, but it’s not significant at the conventional threshold (t value = 1.24).\
- **Any_Fam_Psyc_Hist (0.21)**: Family psychiatric history is positively associated with **QIDS score**, but not statistically significant at the conventional threshold (t value = 1.29).

-   **Random Effects**:
    -   **Intercept Variance (11.81, SD = 3.44)**: There is **substantial variability** in baseline QIDS scores across subjects.\
    -   **Slope Variance for WEEK (0.08, SD = 0.28)**: The rate of change in QIDS scores varies moderately between subjects.
    -   **Residual Variance (10.85, SD = 3.29)**: There is still considerable unexplained variability within subjects.
-   **Correlation of Fixed Effects**:
    -   **Correlation between SEX and Race_Collapsed** is somewhat negative, which may suggest that gender and race have some correlation, but it's **not very strong**.
    -   The **correlation between `dage` and other predictors** is relatively small, indicating little to no collinearity issues.

### **Key Takeaways:**

1.  **WEEK has a strong negative effect** on QIDS scores, meaning participants' depression symptoms decrease over time.
2.  **Race** appears to have a large effect on QIDS scores, with subjects in the **Other** and **Black** categories showing **higher QIDS scores** on average compared to the reference group.
3.  **No meaningful correlation** between baseline QIDS scores and rate of change over time (random intercept and slope correlation is near 0).
4.  **Family psychiatric history (Any_Fam_Psyc_Hist)** does not show a statistically significant effect on QIDS scores.

### **Comparison of Model Fit (lmer.1 vs. lmer.2)**

#### **Model 1 (lmer.1)**:

-   **AIC**: 93,562.5\
-   **BIC**: 93,608.8\
-   **logLik**: -46,775.3\
-   **Deviance**: 93,550.5\
-   **Residual Variance (Std. Dev.)**: 10.85 (3.29)

#### **Model 2 (lmer.2)**:

-   **AIC**: 80,513.3\
-   **BIC**: 80,611.6\
-   **logLik**: -40,243.7\
-   **Deviance**: 80,487.3\
-   **Residual Variance (Std. Dev.)**: 10.75 (3.28)

------------------------------------------------------------------------

### **Interpretation:**

1.  **AIC** and **BIC**:
    -   **lmer.2** has a significantly **lower AIC (80,513.3)** and **BIC (80,611.6)** compared to **lmer.1** (AIC = 93,562.5 and BIC = 93,608.8).\
    -   A lower AIC and BIC indicate **better model fit** in **lmer.2** after including the additional predictors.
2.  **logLik**:
    -   The **log-likelihood** of **lmer.2 (-40,243.7)** is also higher than **lmer.1 (-46,775.3)**, suggesting that the additional predictors in **lmer.2** improve the fit by explaining more variation in the data.
3.  **Residual Variance**:
    -   The residual variance in **lmer.2** (10.75) is slightly lower than **lmer.1** (10.85), suggesting that **lmer.2** accounts for more of the unexplained variance.

### **Conclusion**:

-   The **inclusion of additional predictors (SEX, Race_Collapsed, dage, epino, Any.Fam.Psyc.Hist)** in **lmer.2** improves the model fit, as evidenced by the **lower AIC, BIC**, and **higher log-likelihood**.
-   This suggests that **these predictors** are likely **meaningful** in explaining variation in **QIDS scores** over time.

# c)

```{r}

full_data <- full_data %>%
  mutate(
    dage_scaled = scale(dage),
    epino_scaled = scale(epino),
    episode_date_scaled = scale(episode_date),
    Total_Fam_Psyc_His = DEP + DRUG_PHX + SUIC_PHX + BIP + ALCOHOL
  )

lmer.3 <- lmer(QSTOT ~ 1 + WEEK + Total_Fam_Psyc_His +  
                  dage_scaled + epino_scaled + Race_Collapsed + episode_date_scaled + 
                  (1 + WEEK | ID), 
               data = full_data, 
               REML = F,
               control = lmerControl(optimizer = "Nelder_Mead"))
summary(lmer.3)

```

# Part IV

## a)

```{r}
mynames <- paste("M", as.character(1:3), sep = "")
mymodels <- list(lmer.1, lmer.2, lmer.3)
myaicc <- as.data.frame(aictab(cand.set = mymodels, 
                               modnames = mynames,
                               sort = FALSE)[,-c(5,7)])

myaicc$Eratio <- max(myaicc$AICcWt) / myaicc$AICcWt

data.frame(Modnames = myaicc$Modnames, round(myaicc[,-1], 2))

```

**Comparison of Model Fit:**

**Model 1: lmer.2**
- **AIC**: 80,513.3  
- **BIC**: 80,611.6  
- **logLik**: -40,243.7  
- **Deviance**: 80,487.3  
- **Residual Variance (SD)**: 10.75 (3.28)  

**Model 2: lmer.3**
- **AIC**: 80,240.7  
- **BIC**: 80,338.9  
- **logLik**: -40,107.4  
- **Deviance**: 80,214.7  
- **Residual Variance (SD)**: 10.76 (3.28)

**Key Insights:**

1. **AIC and BIC**:
   - **`lmer.3` has a lower AIC (80,240.7) and BIC (80,338.9)** compared to **`lmer.2` (AIC = 80,513.3, BIC = 80,611.6)**. This suggests that **`lmer.3`** provides a **better fit** to the data, as both AIC and BIC are lower.

2. **logLik**:
   - **`lmer.3` has a higher log-likelihood (-40,107.4)** compared to **`lmer.2` (-40,243.7)**. This indicates that **`lmer.3`** explains more variance in the data.

3. **Residual Variance**:
   - **Residual variance is almost identical** between the two models (10.75 vs. 10.76), suggesting that **`lmer.3`** does not introduce significant extra variability compared to **`lmer.2`**.

---

**Conclusion:**
- **`lmer.3` is the more plausible model** based on the lower AIC, BIC, and higher log-likelihood, suggesting that the inclusion of **`Total_Fam_Psyc_His`** as a composite variable improves the model fit without significantly increasing the residual variance.


1. **Extract Fitted Values for `lmer.3`:**
```{r}

# I was unable to get the fitted values due to missing values so I had to do the following:

full_data_complete <- full_data[complete.cases(full_data[, c("SEX", "Race_Collapsed", "dage_scaled", "epino_scaled", 
                                                             "Total_Fam_Psyc_His", "episode_date_scaled")]), ]

lmer.3 <- lmer(QSTOT ~ 1 + WEEK + Total_Fam_Psyc_His +  
                  dage_scaled + epino_scaled + Race_Collapsed + episode_date_scaled + 
                  (1 + WEEK | ID), 
               data = full_data_complete, 
               REML = F,
               control = lmerControl(optimizer = "Nelder_Mead"))

full_data_complete$fitted_values <- fitted(lmer.3)


```


2. **Select a Random Sample of 25 Subjects:**

```{r}
set.seed(111)
random_sample_ids <- sample(unique(full_data_complete$ID), 25)
subset_data <- full_data_complete %>%
  filter(ID %in% random_sample_ids)

```


3. **Create the Plot:**

```{r}
ggplot(subset_data, aes(x = WEEK, y = QSTOT, group = ID)) +
  geom_point(color = "blue", alpha = 0.5) +  # Raw data points
  geom_line(aes(y = fitted_values), color = "red", size = 1) +  # Predicted values
  facet_wrap(~ID, scales = "free_y") +  # One panel per subject
  labs(title = "Observed vs. Predicted QIDS Scores by Subject",
       x = "Week",
       y = "QIDS Score") +
  theme_minimal()
```


